---
title: "Disposition"
author: "Lia Baumann"
date: "2024-04-17"
output: github_document
bibliography: references.json
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Topic
"Burgundy truffle monitoring"

## Research question(s) and hypotheses
Prelim.:
What is the population structure of the truffles?
Are MLGs perennial?
Differences between or within populations? Interactions... etc?

## Research aims
To learn about the population structure and life cycle of truffles.

## A review of the current state of research
- Truffles in general
- T. aestivum specifically
- results of WSL

## Identification of gaps in the field of research, motivation for the research and justification of the added scientific value for the field of study
- Monitoring data rare, unique

## Materials and Methods
- describe data collection
- describe molecular preparation

## Expected results

## Risk assessment

## Project schedule
- Data collection almost finished
- Data analysis: splitting into different parts

## Budget plan
- Costs: personnel, analyses

## References

The basis for the research about Truffle Monitoring data is described in detail in the following publications:
@molinierFirstIdentificationPolymorphic2013
@molinierSSRbasedIdentificationGenetic2016
@molinierFinescaleGeneticStructure2016
@staubliHiddenFairyRings2022
@steidingerFallSummerTruffle2022
@legendreComparisonMantelTest2010
@kamvarPopprPackageGenetic2014

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
