---
title: "PCA and DAPC"
author: "Lia Baumann"
date: "2024-06-04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = FALSE
)
```

```{r Tabellen laden, message=FALSE, warning=FALSE, include=FALSE}
library(readxl)
library(knitr)
library(tidyverse)
library(ggplot2)
theme_set(theme_bw())
library(envalysis)
library(forcats)
library(poppr)
library(treemap)
library(pegas)
library(adegenet)
library(devtools)
library(hierfstat)
library(mmod)
library(diveRsity)
library(vegan)
library(writexl)
library(devEMF)
library(RColorBrewer)

T_all <- read_excel("Tuaest_Master_genMonit_haploid_MP_Mai2024.xlsx")
myData <- read.genalex("Daten_Genalex.csv",genclone=TRUE, ploidy=1)
myData_genind <- read.genalex("Daten_Genalex.csv",genclone=FALSE, ploidy=1)
myData_genind_allMarkersOnly <- myData_genind %>%
  missingno("geno", cutoff = 0)
splitStrata(myData) <- ~Pop/Month/SamplingYear/TruffleYear
splitStrata(myData_genind) <- ~Pop/Month/SamplingYear/TruffleYear
splitStrata(myData_genind_allMarkersOnly) <- ~Pop/Month/SamplingYear/TruffleYear
cc_myData_genind_allMarkersOnly_SY <- clonecorrect(myData_genind_allMarkersOnly,
                                                   strata=~Pop/SamplingYear)
cc_myData_genind_allMarkersOnly_TY <- clonecorrect(myData_genind_allMarkersOnly,
                                                   strata=~Pop/TruffleYear)
order_sites_WtoE <- factor(c("BUR","GEN","NEU","SCS","KON","FRB","FRE","FRI","SCD","SCG","SCL","WSL","LIM","BRU","RIE","TRO","ALD","UST","BAR","UEB","BOB","BOH","HAN"))

#based on the PCA, we remove BUR, FRB and FRE for some of the analyses and check what it changes
myData_genind_aMO_noOutliers <- popsub(myData_genind_allMarkersOnly,exclude=c("BUR","FRB","FRE"))
cc_myData_genind_aMO_noOutliers <- popsub(cc_myData_genind_allMarkersOnly_SY,exclude=c("BUR","FRB","FRE"))
```


```{r pca test 1, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
from https://www.molecularecologist.com/2015/07/30/pca-of-multilocus-genotypes-in-r/
library(scatterplot3d)
library(rgl)
library(tidyverse)
library(readxl)
library(knitr)
T_all <- read_excel("Tuaest_Master_genMonit_haploid_MP_Mai2024.xlsx")
cc_FullTable_Tuaest_allMarkersOnly <- read.csv("C:/Users/liaba/OneDrive - Eidg. Forschungsanstalt WSL/R/truffles/cc_TUAEST_ALLMARKERSONLY_HWE.csv")
T_all <- separate_wider_delim(T_all,Code_Analyses2, delim="_", names = c("Code","Yr"))
T_all_withMLGs_allMarkersOnly <- inner_join(T_all,cc_FullTable_Tuaest_allMarkersOnly,by=c("Code"="Sample")) %>%
  select(Code, MLG, Site_1_abrev, truffle_year, Sampling_year, Sampling_date, aest06_1.x, aest07_1.x, aest15_1.x, aest26_1.x, aest28_1.x, aest35_1.x, aest36_1.x, aest01_1.x, aest10_1.x, aest18_1.x, aest24_1.x, aest25_1.x, aest29_1.x, aest31_1.x)

T_all_allMarkersOnly.pca<-prcomp(T_all_withMLGs_allMarkersOnly[,7:20], center=TRUE,scale.=TRUE)
pcs123<-T_all_allMarkersOnly.pca$x[,1:3]
WSL<-grep("WSL",T_all_withMLGs_allMarkersOnly$Site_1_abrev)
BUR<-grep("BUR",T_all_withMLGs_allMarkersOnly$Site_1_abrev)
BRU<-grep("BRU",T_all_withMLGs_allMarkersOnly$Site_1_abrev)
SCG<-grep("SCG",T_all_withMLGs_allMarkersOnly$Site_1_abrev)

p<-scatterplot3d(pcs123[WSL,1],pcs123[WSL,2],pcs123[WSL,3],color=c("red"),
                 pch=20,xlab="PC1",ylab="PC2",zlab="PC3",
                 xlim=c(min(pcs123[,1]),max(pcs123[,1])),
                 ylim=c(min(pcs123[,2]),max(pcs123[,2])),
                 zlim=c(min(pcs123[,3]),max(pcs123[,3])))

p$points3d(pcs123[BUR,1],pcs123[BUR,2], pcs123[BUR,3],col=c("blue"),pch=20)
p$points3d(pcs123[BRU,1],pcs123[BRU,2], pcs123[BRU,3],col=c("maroon"),pch=20)
p$points3d(pcs123[SCG,1],pcs123[SCG,2], pcs123[SCG,3],col=c("green"),pch=20)
legend(p$xyz.convert(-80, 20, -10), col= c("red","blue","maroon","green"), bg="white", pch=c(20,20,20,20), yjust=0,
  legend = c("WSL", "BUR", "BRU", "SCG"), cex = 1.1)
```

## 1. PCA

```{r pca, echo=TRUE}
#uncorrected PCA with allMarkersOnly dataset
setPop(myData_genind_allMarkersOnly) <- ~Pop
x.pops_allMarkersOnly <- tab(myData_genind_allMarkersOnly,
                             freq=TRUE, NA.method="mean")

pca.pops.allMarkersOnly <- dudi.pca(x.pops_allMarkersOnly,center=TRUE,scale=FALSE, nf=3, scannf=FALSE)
s.label(pca.pops.allMarkersOnly$li)
#looking at first two axes:
s.class(pca.pops.allMarkersOnly$li,fac=pop(myData_genind_allMarkersOnly),col=transp(funky(15),0.6),axesel=FALSE,cstar=0,cpoint=3)
add.scatter.eig(pca.pops.allMarkersOnly$eig[1:50],3,1,2, ratio=.17)
#looking at 2 and 3 axis:
s.class(pca.pops.allMarkersOnly$li, fac=pop(myData_genind_allMarkersOnly),
xax=2, yax=3, col=transp(funky(15),.6),
axesel=FALSE, cstar=0, cpoint=3)
add.scatter.eig(pca.pops.allMarkersOnly$eig[1:50],3,2,3, ratio=.2)


#emf(file="pca.allMarkersOnly.emf")
#s.class(pca.pops.allMarkersOnly$li, fac=pop(myData_genind_allMarkersOnly),
 #       col=funky(15))
#dev.off()

#check after removal of outliers:
setPop(myData_genind_aMO_noOutliers) <- ~Pop
x.pops_allMarkersOnly_noOutliers <- tab(myData_genind_aMO_noOutliers,
                             freq=TRUE, NA.method="mean")

pca.pops.aMO.noOutliers <- dudi.pca(x.pops_allMarkersOnly_noOutliers,center=TRUE,scale=FALSE, nf=3, scannf=FALSE)
s.label(pca.pops.aMO.noOutliers$li)
#looking at first two axes:
s.class(pca.pops.aMO.noOutliers$li,fac=pop(myData_genind_aMO_noOutliers),col=transp(funky(15),0.6),axesel=FALSE,cstar=0,cpoint=3)
add.scatter.eig(pca.pops.aMO.noOutliers$eig[1:50],3,1,2, ratio=.17)
#looking at 2 and 3 axis:
s.class(pca.pops.aMO.noOutliers$li, fac=pop(myData_genind_aMO_noOutliers),
xax=2, yax=3, col=transp(funky(15),.6),
axesel=FALSE, cstar=0, cpoint=3)
add.scatter.eig(pca.pops.aMO.noOutliers$eig[1:50],3,2,3, ratio=.2)


#clonecorrected PCA  with samplingYear
setPop(cc_myData_genind_allMarkersOnly_SY) <- ~Pop
x.pops_cc.SY <- tab(cc_myData_genind_allMarkersOnly_SY,
                    freq=TRUE, NA.method="mean")
pca.pops_cc.SY <- dudi.pca(df = x.pops_cc.SY,
                           center = TRUE, scale = FALSE, scannf = FALSE, nf = 3)
#looking at first two axes:
s.class(pca.pops_cc.SY$li,
        fac=pop(cc_myData_genind_allMarkersOnly_SY),
        col=transp(funky(15),0.6),axesel=FALSE,cstar=0,cpoint=3)
add.scatter.eig(pca.pops_cc.SY$eig[1:50],3,1,2, ratio=.17)

#looking at 2 and 3 axis:
s.class(pca.pops_cc.SY$li, fac=pop(cc_myData_genind_allMarkersOnly_SY),
xax=2, yax=3, col=transp(funky(15),.6),
axesel=FALSE, cstar=0, cpoint=3)
add.scatter.eig(pca.pops_cc.SY$eig[1:50],3,2,3, ratio=.2)

#emf(file="pca.cc.allMarkersOnly_SY.emf")
#s.class(pca.pops_cc.SY$li,
#        fac=pop(cc_myData_genind_allMarkersOnly_SY),
#        col=funky(15))
#dev.off()

#check after removal of outliers:
setPop(cc_myData_genind_aMO_noOutliers) <- ~Pop
x.pops.cc_allMarkersOnly_noOutliers <- tab(cc_myData_genind_aMO_noOutliers,
                             freq=TRUE, NA.method="mean")

pca.pops.cc.aMO.noOutliers <- dudi.pca(x.pops.cc_allMarkersOnly_noOutliers,center=TRUE,scale=FALSE, nf=3, scannf=FALSE)
s.label(pca.pops.cc.aMO.noOutliers$li)
#looking at first two axes:
s.class(pca.pops.cc.aMO.noOutliers$li,fac=pop(cc_myData_genind_aMO_noOutliers),col=transp(funky(15),0.6),axesel=FALSE,cstar=0,cpoint=3)
add.scatter.eig(pca.pops.cc.aMO.noOutliers$eig[1:50],3,1,2, ratio=.17)
#looking at 2 and 3 axis:
s.class(pca.pops.cc.aMO.noOutliers$li, fac=pop(cc_myData_genind_aMO_noOutliers),
xax=2, yax=3, col=transp(funky(15),.6),
axesel=FALSE, cstar=0, cpoint=3)
add.scatter.eig(pca.pops.cc.aMO.noOutliers$eig[1:50],3,2,3, ratio=.2)

#clonecorrected PCA  with TruffleYear
setPop(cc_myData_genind_allMarkersOnly_TY) <- ~Pop
x.pops_cc.TY <- tab(cc_myData_genind_allMarkersOnly_TY,
                    freq=TRUE, NA.method="mean")
pca.pops_cc.TY <- dudi.pca(df = x.pops_cc.TY,
                           center = TRUE, scale = FALSE, scannf = FALSE, nf = 3)
#looking at first two axes:
s.class(pca.pops_cc.TY$li,
        fac=pop(cc_myData_genind_allMarkersOnly_TY),
        col=transp(funky(15),0.6),axesel=FALSE,cstar=0,cpoint=3)
add.scatter.eig(pca.pops_cc.TY$eig[1:50],3,1,2, ratio=.17)

#looking at 2 and 3 axis:
s.class(pca.pops_cc.TY$li, fac=pop(cc_myData_genind_allMarkersOnly_TY),
xax=2, yax=3, col=transp(funky(15),.6),
axesel=FALSE, cstar=0, cpoint=3)
add.scatter.eig(pca.pops_cc.TY$eig[1:50],3,2,3, ratio=.2)

#emf(file="pca.cc.allMarkersOnly_TY.emf")
#s.class(pca.pops_cc.TY_2$li,
#        fac=pop(cc_myData_genind_allMarkersOnly_TY),
#        col=funky(15))
#dev.off()

#plot(pca.pops_cc.TY$li, col=
```

```{r another pca}
setPop(myData_genind_allMarkersOnly) <- ~Pop
pca1 = dudi.pca(myData_genind_allMarkersOnly, scannf = FALSE, nf = 3)
cols = brewer.pal(nPop(myData_genind_allMarkersOnly), "RdYlBu")

# Percent of genetic variance explained by each axis
percent = round(pca1$eig/sum(pca1$eig)*100, digits = 1)
scatter1 <- scatter_plot(
  dataframe = pca1$li,
  group_ids = myData_genind_allMarkersOnly$pop,
  type = "points",
  axes = c(1,2),
  percent = percent,
  colours = cols,
  point_size = 2,
  point_type = 21,
  centroid_size = 2,
  stroke = 0.1,
  plot_title = "PCA coloured by group_ids"
)+
  theme(
    legend.position = "none",
    axis.title = element_text(size = 8),
    axis.text = element_text(size = 6),
    plot.title = element_text(size = 10),
  )
scatter1
```
# 2. DAPC

DAPC was pioneered by Jombart and colleagues (Jombart et al., 2010) and can be used to infer the number of clusters of genetically related individuals. In this multivariate statistical approach variance in the sample is partitioned into a between-group and within- group component, in an effort to maximize discrimination between groups. In DAPC, data is first transformed using a principal components analysis (PCA) and subsequently clusters are identified using discriminant analysis (DA). This tutorial is based on the vignette written by Thibaut Jombart. We encourage the user to explore this vignette further. The vignette can also be opened within R by executing adegenetTutorial("dapc")

```{r dapc}
setPop(myData_genind_allMarkersOnly) <- ~Pop
dapc.allMarkers <- dapc(myData_genind_allMarkersOnly, var.contrib = TRUE, scale = FALSE, n.pca = 30, n.da = nPop(myData_genind_allMarkersOnly) - 1)
scatter(dapc.allMarkers, cell = 0, pch = 18:23, cstar = 0, mstree = TRUE, lwd = 2, lty = 2)

#clonecorrected


#BUR, FRB and FRE are separated from the cluster by their second axis (y) --> genetically distinct
#look at the DAPC without these populations:
setPop(myData_genind_aMO_noOutliers) <- ~Pop
dapc.allMarkers_noOutliers <- dapc(myData_genind_aMO_noOutliers, var.contrib = TRUE, scale = FALSE, n.pca = 30, n.da = nPop(myData_genind_aMO_noOutliers) - 1)
scatter(dapc.allMarkers_noOutliers, cell = 0, pch = 18:23, cstar = 0, mstree = TRUE, lwd = 2, lty = 2)

#and clonecorrected without outliers:
setPop(cc_myData_genind_aMO_noOutliers) <- ~Pop
dapc.cc.allMarkers_noOutliers <- dapc(cc_myData_genind_aMO_noOutliers, var.contrib = TRUE, scale = FALSE, n.pca = 30, n.da = nPop(cc_myData_genind_aMO_noOutliers) - 1)
scatter(dapc.cc.allMarkers_noOutliers, cell = 0, pch = 18:23, cstar = 0, mstree = TRUE, lwd = 2, lty = 2)

```