---
title: "Persistence of MLGs with all markers"
author: "Lia Baumann"
date: "2024-05-29"
output: 
  github_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Tabellen laden, message=FALSE, warning=FALSE, include=FALSE}
library(readxl)
library(knitr)
library(tidyverse)
library(ggplot2)
theme_set(theme_bw())
library(sf)
library(ggspatial)
library(rnaturalearth)
library(envalysis)
library(forcats)
library(poppr)
library(treemap)
library(pegas)
library(adegenet)
library(devtools)
library(hierfstat)
library(mmod)
library(scales)
library(plotly)
library(devEMF)
T_all <- read_excel("Tuaest_Master_genMonit_haploid_MP_Mai2024.xlsx")
# wirf alle Individuals mit NA raus
myData <- read.genalex("Daten_Genalex.csv",genclone=TRUE, ploidy=1)
splitStrata(myData) <- ~Pop/Month/SamplingYear/TruffleYear
setPop(myData) <- ~Pop
Tuaest_allMarkersOnly <- myData %>%
  missingno("geno", cutoff = 0)
cc_FullTable_Tuaest_allMarkersOnly <- read.csv("C:/Users/liaba/OneDrive - Eidg. Forschungsanstalt WSL/R/truffles/cc_TUAEST_ALLMARKERSONLY_HWE.csv")
cc_distinctMLGs_allMarkersOnly <- cc_FullTable_Tuaest_allMarkersOnly[!duplicated(cc_FullTable_Tuaest_allMarkersOnly$MLG), ]
T_all <- separate_wider_delim(T_all,Code_Analyses2, delim="_", names = c("Code","Yr"))
T_all_withMLGs_allMarkersOnly <- inner_join(T_all,cc_FullTable_Tuaest_allMarkersOnly,by=c("Code"="Sample")) %>%
  select(Code, MLG, Site_1_abrev, truffle_year, Sampling_year,Sampling_date)

clusters <- read_excel("STRUCTURE Results MLGs.xlsx")
clusters$`clusterssign(q>0.5)` <- factor(clusters$`clusterssign(q>0.5)`)
T_all_withMLGs_withClusters_allMarkersOnly <- left_join(T_all_withMLGs_allMarkersOnly,clusters,by=c("MLG"="MLG_ID")) %>% select(-"IndID2") %>% distinct(.)
```

## Genetic information

Erklärung der Datenherkunft:

1.  myData = Daten_Genalex.csv

Aus myData werden danach alle missing (NA) gefiltert:

Tuaest_allMarkersOnly \<- myData %\>% missingno("geno", cutoff = 0)

```{r poppr with Tuaest_allMarkersOnly, eval=FALSE, include=FALSE}
setPop(Tuaest_allMarkersOnly) <- ~Pop
poppr_withoutNA <- poppr(Tuaest_allMarkersOnly)
poppr_withoutNA_genind <- poppr(myData_genind_allMarkersOnly)
kable(select(poppr_withoutNA,Pop,N,MLG,eMLG,SE,H,G,lambda,E.5,Hexp,Ia,rbarD))
kable(select(poppr_withoutNA_genind, Pop, N, MLG, eMLG, SE, H, G, lambda, E.5, Hexp, Ia, rbarD))

# sie machen dasselbe! die Klonkorrektur kann direkt über Poppr erfolgen :)
```

Info zur Tabelle: N = Number of individuals, MLG = Number of multilocus
genotypes, eMLG = number of expected MLG at the smallest sample size \>=
10 based on   rarefaction SE = Standard error based on eMLG, H =
Shannon-Wiener Index of MLG diversity G = Stoddart & Taylor's Index of
MLG diversity lambda = Simpsons index, E.5 = Evenness, Hexp = Neis
Expected Heterozygosity Ia = Index of association, rbarD = stand. Index
of association

-- Wegen der wenigen Samples habe ich SCHIF und BAR danach ausgelassen.
–

Hier nochmals die gleiche Tabelle, jedoch ohne Populationen mit kleiner Sample Size (n<20):
```{r poppr with Tuaest_allMarkersOnly only sites with 20+ samples, eval=FALSE, include=FALSE}
setPop(Tuaest_allMarkersOnly) <- ~Pop
poppr_withoutNA_min20samples <- poppr(popsub(Tuaest_allMarkersOnly, exclude=c("SCG","FRB","BAR","GEN","SCHIF")))
kable(select(poppr_withoutNA_min20samples,Pop,N,MLG,eMLG,SE,H,G,lambda,E.5,Hexp,Ia,rbarD))
```

## Persistence over the years (MLG per population)

1.  Import der klonkorrigierten Datei "cc_TUAEST_ALLMARKERSONLY", welche
    mit MLGSim generiert wurde (enthält Sample, MLG, PSex sowie die
    Marker):

cc_FullTable_Tuaest_allMarkersOnly \<-
read.csv("C:/Users/liaba/OneDrive - Eidg. Forschungsanstalt
WSL/R/truffles/cc_TUAEST_ALLMARKERSONLY_HWE.csv")

2.  MLG Duplikate werden entfernt

cc_distinctMLGs_allMarkersOnly \<-
cc_FullTable_Tuaest_allMarkersOnly[!duplicated(cc_FullTable_Tuaest_allMarkersOnly\$MLG),]

3.  T_all kommt aus der Masterliste
    Tuaest_Master_genMonit_haploid_MP_Mai2024.xlsx

Hier wird zuerst die Spalte "Code_Analyses2" vom Jahr separiert, und
danach verjoint, damit die MLGs zusammen mit den Sample-Informationen
verfügbar sind:

T_all \<- separate_wider_delim(T_all,Code_Analyses2, delim="\_", names =
c("Code","Yr"))

T_all_withMLGs_allMarkersOnly \<-
inner_join(T_all,cc_FullTable_Tuaest_allMarkersOnly,by=c("Code"="Sample"))
%\>% select(Code, MLG, Site_1_abrev, truffle_year,
Sampling_year,Sampling_date)

4.  Nun können wir damit die Grafiken generieren:

## Looking at persistence over years


--\> bei LIM und BRU kommen die MLGs doppelt vor bzw. an beiden Orten...
sollte man diese vielleicht zusammen nehmen?

```{r some persistence stats}
# Count the number of unique years each individual was found
year_count_k6 <- T_all_withMLGs_withClusters_allMarkersOnly %>%
  group_by(MLG) %>%
  summarize(years_found = n_distinct(Sampling_year))
year_count_k6

year_mean_k6 <- year_count_k6 %>%
  summarize(mean = mean(years_found))
year_mean_k6

# Find the maximum number of years any MLG has been found
max_years <- max(year_count_k6$years_found)
# Filter the data to include only MLGs found for the maximum number of years
mlgs_longest_found <- subset(year_count_k6, years_found == max_years)
mlgs_longest_found

proportion <- year_count_k6 %>%
  mutate(found_once = ifelse(years_found == 1, "Once", "Multiple")) %>%
  count(found_once) %>%
  mutate(proportion = n / sum(n))
proportion

# Join back to the original data to classify each sample as "Once" or "Multiple"
MLGcount_with_classification <- T_all_withMLGs_withClusters_allMarkersOnly %>%
  left_join(year_count_k6, by = "MLG") %>%
  mutate(found_once = ifelse(years_found == 1, "Once", "Multiple")) %>%
  count(found_once) %>%
  mutate(proportion = n / sum(n))
MLGcount_with_classification #over all samples

# Count the number of samples for each individual
sample_count <- T_all_withMLGs_withClusters_allMarkersOnly %>%
  group_by(MLG) %>%
  summarize(total_samples = n())
#write.csv(sample_count, "powerful_mlgs.csv")

sample_count <- sample_count %>%
  mutate(proportion = total_samples / sum(total_samples))

# Count how many individuals produced each sample count
individuals_per_sample_count <- sample_count %>%
  count(total_samples, name = "number_of_individuals") %>%
  mutate(total_samples_contributed = total_samples * number_of_individuals,
         proportion_of_total_samples = total_samples_contributed / sum(total_samples_contributed)*100)
individuals_per_sample_count

#plot pie chart
top_samples <- sample_count %>%
  arrange(desc(total_samples)) %>%
  mutate(MLG = ifelse(row_number() <= 10, as.character(MLG), "Other")) %>%
  group_by(MLG) %>%
  summarize(total_samples = sum(total_samples))

# Calculate the proportions
top_samples <- top_samples %>%
  mutate(proportion = total_samples / sum(total_samples))

# Generate a color palette with RColorBrewer
num_colors <- n_distinct(top_samples$MLG)  # Count of unique individuals
colors <- colorspace::sequential_hcl(num_colors)  # Choose a palette (e.g., "Blues")

# Create a pie chart
ggplot(top_samples, aes(x = "", y = proportion, fill = MLG)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Proportion of Samples by MLG (Top 10 + Others)",
       fill = "MLG") +
  scale_fill_manual(values = colors) +
  theme_void() +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 8))

mlg_data <- data.frame(
  MLG = c("MLG1", "MLG2", "MLG3", "MLG4", "MLG5", "MLG6"),
  YearsFound = c(3, 5, 5, 2, 4, 3),
  Cluster = c("Cluster1", "Cluster1", "Cluster2", "Cluster2", "Cluster3", "Cluster3")
)

mlg_cluster_test <- year_count_k6 %>%
  inner_join(T_all_withMLGs_withClusters_allMarkersOnly,by="MLG")

#mlg_cluster_test_glm <- mlg_cluster_test %>%
#  rename(Cluster='clusterssign(q>0.5)') %>%
#  select(MLG,years_found,Cluster)
#mlg_cluster_test_glm1 <- mlg_cluster_test_glm %>%
#  group_by(MLG) %>%
#  summarize(SampleSize=n())%>%
 # ungroup()
#glm <- left_join(mlg_cluster_test_glm1,mlg_cluster_test_glm,by="MLG") %>%
#  group_by(MLG) %>%
#  distinct(MLG) %>%
#  ungroup()

# Summarize data by Cluster
# Calculate mean perenniality and sample size per cluster
cluster_summary <- mlg_cluster_test %>%
  group_by(Cluster=`clusterssign(q>0.5)`) %>%
  summarize(
    MeanYearsFound = mean(years_found),
    SampleSize = n()
  )
cluster_summary


# Fit a GLM to check if there's a significant difference in perenniality across clusters
# Correct for sample size by including SampleSize as a covariate
glm_model <- glm(MeanYearsFound ~ Cluster + SampleSize, data=glm, family = poisson(link="log"))

# Summary of the model to assess significance of Cluster effect
summary(glm_model)
```

```{r MLGs per year per surface}
persistence_bob <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "BOB"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="distribution of MLGs in BOB (Bohlingen Buche) over the years") + scale_x_continuous(breaks=2010:2023)
persistence_bob
```
```{r emf, eval=FALSE, include=FALSE}
emf(file="persistence_bob.emf")
persistence_bob
dev.off()

```

```{r more MLGs without code display, echo=FALSE}
persistence_ald <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "ALD"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in ALD over the years") + scale_x_continuous(breaks=2010:2023)
persistence_ald
emf(file="persistence_ald.emf")
persistence_ald
dev.off()

persistence_bru <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "BRU"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in BRU over the years") + scale_x_continuous(breaks=2010:2024)
persistence_bru
emf(file="persistence_bru.emf")
persistence_bru
dev.off()

persistence_bur <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "BUR"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in BUR over the years") + scale_x_continuous(breaks=2010:2023)
persistence_bur
emf(file="persistence_bur.emf")
persistence_bur
dev.off()

persistence_frb <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "FRB"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in FRB over the years") + scale_x_continuous(breaks=2010:2023)
persistence_frb
emf(file="persistence_frb.emf")
persistence_frb
dev.off()

persistence_fre <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "FRE"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in FRE over the years") + scale_x_continuous(breaks=2010:2023)
persistence_fre
emf(file="persistence_fre.emf")
persistence_fre
dev.off()

persistence_fri <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "FRI"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in FRI over the years") + scale_x_continuous(breaks=2010:2023)
persistence_fri
emf(file="persistence_fri.emf")
persistence_fri
dev.off()

ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "GEN"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in GEN over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "HAN"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in HAN over the years") + scale_x_continuous(breaks=2010:2023)
persistence_kon <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "KON"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in KON over the years") + scale_x_continuous(breaks=2010:2024)

persistence_lim <- ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "LIM"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in LIM over the years") + scale_x_continuous(breaks=2010:2024)
persistence_lim
emf(file="persistence_lim.emf")
persistence_lim
dev.off()

ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "NEU"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in NEU over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "RIE"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in RIE over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "SCD"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in SCD over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "SCG"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in SCG over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "SCL"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in SCL over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "SCS"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in SCS over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "TRO"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in TRO over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "UEB"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in UEB over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "UST"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in UST over the years") + scale_x_continuous(breaks=2010:2023)
ggplot(subset(T_all_withMLGs_allMarkersOnly, Site_1_abrev %in% "WSL"),aes(Sampling_year,MLG)) + geom_point() + geom_line(aes(group=MLG)) +
  labs(title ="Distribution of MLGs in WSL over the years") + scale_x_continuous(breaks=2010:2023)

#emf(file="persistence_kon.emf")
#persistence_kon
#dev.off()
```
