---
title: "Richness, Evenness, AMOVA, Isolation by distance"
author: "Lia Baumann"
date: "2024-09-03"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r Tabellen laden, message=FALSE, warning=FALSE, include=FALSE}
library(readxl)
library(knitr)
library(tidyverse)
library(ggplot2)
theme_set(theme_bw())
library(envalysis)
library(forcats)
library(poppr)
library(treemap)
library(pegas)
library(adegenet)
library(devtools)
library(hierfstat)
library(mmod)
library(diveRsity)
library(vegan)
library(writexl)
library(devEMF)
library(ape)
library(magrittr)
library(reshape)
library(reshape2)
library(graph4lg)
library(iNEXT)
#library(LandGenCourse) 
library(GeNetIt)
library(EcoGenetics)
library(gstudio)
#library(PopGenReport)
#library(spdep)
library(popgraph)

T_all <- read_excel("Tuaest_Master_genMonit_haploid_MP_Mai2024.xlsx")
coord <- read_excel("Koordinaten_2024.xlsx")
myData <- read.genalex("Daten_Genalex.csv",genclone=TRUE, ploidy=1)
myData_genind <- read.genalex("Daten_Genalex.csv",genclone=FALSE, ploidy=1)
myData_genind_allMarkersOnly <- myData_genind %>%
  missingno("geno", cutoff = 0)
splitStrata(myData) <- ~Pop/Month/SamplingYear/TruffleYear
splitStrata(myData_genind) <- ~Pop/Month/SamplingYear/TruffleYear
splitStrata(myData_genind_allMarkersOnly) <- ~Pop/Month/SamplingYear/TruffleYear
cc_myData_genind_allMarkersOnly_SY <- clonecorrect(myData_genind_allMarkersOnly,
                                                   strata=~Pop/SamplingYear)
cc_myData_genind_allMarkersOnly_TY <- clonecorrect(myData_genind_allMarkersOnly,
                                                   strata=~Pop/TruffleYear)

myData_genind_withClusters <- read.genalex("DatenGenalex_withMLGs_withClusters_allMarkersOnly.csv",genclone=FALSE,ploidy=1)
splitStrata(myData_genind_withClusters) <- ~Pop/Month/SamplingYear/TruffleYear/cluster
```


```{r comparing datareduction}

# Assuming `genind_original` and `genind_reduced` are your two genind objects
datasets <- list(Original = myData_genind, Reduced = myData_genind_allMarkersOnly)

#function to calculate metrics
calculate_metrics <- function(genind_obj) {
  # Number of individuals
  num_individuals <- nInd(genind_obj)
  
  # Number of loci
  num_loci <- nLoc(genind_obj)
  
  #Number of alleles
  # Calculate alleles per locus
  alleles_per_locus <- apply(tab(genind_obj), 2, function(x) length(unique(x[!is.na(x)])))
  total_num_alleles <-  sum(alleles_per_locus, na.rm = TRUE)
  
  # Calculate allelic richness manually
  alleles_per_locus <- apply(tab(genind_obj), 2, function(x) length(unique(x[!is.na(x)])))
  allelic_richness <- mean(alleles_per_locus, na.rm = TRUE)
  
  # Genetic diversity (Nei's diversity)
  nei_diversity <- mean(Hs(genind_obj), na.rm = TRUE)
  
  # Missing data proportion
  missing_data_prop <- sum(is.na(tab(genind_obj))) / length(tab(genind_obj))
  
  
  # Extract the genotype data from the genind object
  genotypes <- as.data.frame(genind_obj@tab)

  # Create a unique identifier for each MLG by combining allele information across all loci
  # You may need to adjust the indices based on your specific data structure
  genotypes$MLG <- apply(genotypes, 1, function(x) paste(na.omit(x), collapse = "_"))

  # Count the number of unique MLGs
  num_distinct_MLGs <- length(unique(genotypes$MLG))

  return(data.frame(
    "Number_of_Individuals" = num_individuals,
    "Number_of_Loci" = num_loci,
    "Number_of_Alleles" = total_num_alleles,
    "Allelic_Richness" = allelic_richness,
    "Number_of_distinct_MLGs" = num_distinct_MLGs,
    "Nei_Diversity" = nei_diversity,
    "Missing_Data_Proportion" = missing_data_prop
  ))
}

# Apply the function to both datasets and combine results
metrics_df <- do.call(rbind, lapply(datasets, calculate_metrics))
metrics_df <- cbind(Dataset = rownames(metrics_df), metrics_df)
rownames(metrics_df) <- NULL

# Display the metrics
print(metrics_df)

# Reshape metrics_df to long format and define the order of metrics
metrics_long <- metrics_df %>%
  pivot_longer(cols = -Dataset, names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = factor(Metric, levels = c(
    "Number_of_Individuals",
    "Number_of_Loci",
    "Number_of_Alleles",
    "Allelic_Richness",
    "Number_of_distinct_MLGs",
    "Nei_Diversity",
    "Missing_Data_Proportion"
  )))
metrics_long

# Plot with improved visibility and specified metric order
ggplot(metrics_long, aes(x = Metric, y = Value, fill = Dataset)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.7), color = "black") +
  geom_text(aes(label = round(Value, 2)), 
            position = position_dodge(width = 0.7), 
            vjust = -0.5, 
            size = 4.5) +  # Adjust text size
  scale_fill_manual(values = c("Original" = "skyblue", "Reduced" = "salmon")) +
  labs(title = "Comparison of Original and Reduced Datasets",
       y = "",
       x = "") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  theme_minimal(base_size = 14) +  # Increase base font size
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better visibility


```


```{r crosspop mlgs}
mlg_data <- data.frame(
  mlg = mlg.vector(myData_genind_allMarkersOnly),
  pop = pop(myData_genind_allMarkersOnly)
)
shared_mlgs <- mlg_data %>%
  group_by(mlg) %>%
  summarize(n_pops = n_distinct(pop)) %>%
  filter(n_pops > 1)
print(shared_mlgs)
```

# Hardy-Weinberg equilibrium
Cannot be determined for haploids.

# Genotypic richness
The number of observed MLGs is equivalent to genotypic richness. We expect that the BB population would have a higher genotypic richness than the FR population. However, looking at the raw number of MLGs for each symptom type, it shows us the opposite: there are 94 MLGs for BB and 191 MLGs for FR. This discrepancy has to do with the sample size differences, namely N=113 for BB and N=581 for FR. A more appropriate comparison is the eMLG value, which is an approximation of the number of genotypes that would be expected at the largest, shared sample size (N=113) based on rarefaction. For BB (N=113) the eMLG=94 and for FR (where N is set to 113) the eMLG = 66.6. Thus, genotypic richness is indeed higher in the BB populations than the FR population when considering equal sample sizes.

```{r richness}

setPop(myData_genind_allMarkersOnly) <- ~SamplingYear
# Calculating minimal necessary sample size
# Step 1: Extract population information

# Step 1: Extract population information
populations <- pop(myData_genind_allMarkersOnly)

# Step 2: Calculate sample sizes per population
population_counts <- table(populations)

# Step 3: Calculate number of alleles per locus
allele_counts <- nAll(myData_genind_allMarkersOnly)

# Step 4: Create a data frame to hold alleles and populations
# Repeat population assignments for each locus
pop_assignments <- rep(populations, each = length(allele_counts))

# Combine the population assignments and allele counts into a data frame
allele_data <- data.frame(Population = pop_assignments,
                           Locus = rep(1:length(allele_counts), times = length(populations)),
                           AlleleCount = rep(allele_counts, each = length(populations)))

# Step 5: Calculate average number of alleles per population
average_alleles <- aggregate(AlleleCount ~ Population, data = allele_data, FUN = mean)

# Combine sample sizes and average alleles into one data frame
sample_summary <- data.frame(Population = names(population_counts),
                             SampleSize = as.integer(population_counts),
                             AverageAlleles = average_alleles$AlleleCount)

# Step 6: Plot Sample Size vs. Average Number of Alleles
ggplot(sample_summary, aes(x = SampleSize, y = AverageAlleles)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Sample Size vs. Average Number of Alleles (Stratum SamplingYear)",
       x = "Sample Size",
       y = "Average Number of Alleles") +
  theme_minimal()

#Minimal sample size Sum is 1 so this can not give any relevant output. I will plot the same for higher sample sizes only
#only calculate Richness for populations > 100 samples (N)
myData_genind_allMarkersOnly_above100N <- selPopSize(myData_genind_allMarkersOnly,pop=NULL,nMin=100)
summary(myData_genind_allMarkersOnly_above100N)
H.year <- mlg.table(myData_genind_allMarkersOnly_above100N, plot = FALSE)
rarecurve(H.year, ylab="Number of expected MLGs", sample=min(rowSums(H.year)),
          border = NA, fill = NA, font = 2, cex = 1, col = "blue")
title("Rarecurve of Sampling Years (min. sample size = 100 per year)")



setPop(myData_genind_allMarkersOnly) <- ~Pop
# Calculating minimal necessary sample size
# Step 1: Extract population information

# Step 1: Extract population information
populations <- pop(myData_genind_allMarkersOnly)

# Step 2: Calculate sample sizes per population
population_counts <- table(populations)

# Step 3: Calculate number of alleles per locus
allele_counts <- nAll(myData_genind_allMarkersOnly)

# Step 4: Create a data frame to hold alleles and populations
# Repeat population assignments for each locus
pop_assignments <- rep(populations, each = length(allele_counts))

# Combine the population assignments and allele counts into a data frame
allele_data <- data.frame(Population = pop_assignments,
                           Locus = rep(1:length(allele_counts), times = length(populations)),
                           AlleleCount = rep(allele_counts, each = length(populations)))

# Step 5: Calculate average number of alleles per population
average_alleles <- aggregate(AlleleCount ~ Population, data = allele_data, FUN = mean)

# Combine sample sizes and average alleles into one data frame
sample_summary <- data.frame(Population = names(population_counts),
                             SampleSize = as.integer(population_counts),
                             AverageAlleles = average_alleles$AlleleCount)

# Step 6: Plot Sample Size vs. Average Number of Alleles
ggplot(sample_summary, aes(x = SampleSize, y = AverageAlleles)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Sample Size vs. Average Number of Alleles (Stratum Pop)",
       x = "Sample Size",
       y = "Average Number of Alleles") +
  theme_minimal()

setPop(myData_genind_allMarkersOnly) <- ~Pop
myData_genind_allMarkersOnly_above5N <- selPopSize(myData_genind_allMarkersOnly,pop=NULL,nMin=5)
poppr(myData_genind_allMarkersOnly_above5N)
H.year <- mlg.table(myData_genind_allMarkersOnly_above5N, plot = FALSE)
rarecurve(H.year, ylab="Number of expected MLGs", sample=min(rowSums(H.year)),
          border = NA, fill = NA, font = 2, cex = 1, col = "blue")
title("Rarecurve of Pop (min. Sample size = 5 per Pop)")

#rescale to see anything except for BOB
setPop(myData_genind_allMarkersOnly) <- ~Pop
myData_genind_allMarkersOnly_above5N_withoutBOB <- popsub(myData_genind_allMarkersOnly_above5N,exclude="BOB")
poppr(myData_genind_allMarkersOnly_above5N_withoutBOB)
H.year_withoutBOB <- mlg.table(myData_genind_allMarkersOnly_above5N_withoutBOB, plot = FALSE)
rarecurve(H.year_withoutBOB, ylab="Number of expected MLGs", sample=min(rowSums(H.year_withoutBOB)),
font = 2, cex = 1, col = "blue")
title("Rarecurve of Pop  (min. Sample size = 5 per Pop; without BOB)")


setPop(myData_genind_allMarkersOnly) <- ~TruffleYear
myData_genind_allMarkersOnly_above100N <- selPopSize(myData_genind_allMarkersOnly,pop=NULL,nMin=100)
summary(myData_genind_allMarkersOnly_above100N)
H.year <- mlg.table(myData_genind_allMarkersOnly_above100N, plot = FALSE)
rarecurve(H.year, ylab="Number of expected MLGs", sample=min(rowSums(H.year)),
          border = NA, fill = NA, font = 2, cex = 1, col = "blue")
title("Rarecurve of Truffle Years")

setPop(myData_genind_allMarkersOnly) <- ~Month
myData_genind_allMarkersOnly_above40N <- selPopSize(myData_genind_allMarkersOnly,pop=NULL,nMin=40)
summary(myData_genind_allMarkersOnly_above40N)
poppr(myData_genind_allMarkersOnly_above40N)
H.year <- mlg.table(myData_genind_allMarkersOnly_above40N, plot = FALSE)
rarecurve(H.year, ylab="Number of expected MLGs", sample=min(rowSums(H.year)),
          border = NA, fill = NA, font = 2, cex = 1, col = "blue")
title("Rarecurve of Month")



#Genotypic richness rises, the later the ascocarps were collected --> more different individuals produced fruiting bodies
```
# Population structure using Fst

```{r Fst values}
truffles.matFst <- mat_pw_fst(myData_genind_allMarkersOnly)
truffles.matFst
truffles.tree <- nj(truffles.matFst)
plot(truffles.tree, type="unr", tip.col=funky(nPop(myData_genind_allMarkersOnly)), font=2)
annot <- round(truffles.tree$edge.length,2)
edgelabels(annot[annot>0], which(annot>0), frame="n")
add.scale.bar()

table.paint(truffles.matFst,col.labels=1:23)

boxplot(truffles.matFst, col=funky(nPop(myData_genind_allMarkersOnly)), las=3,
xlab="Population", ylab="Fst")

```
# Population structure using Gst

```{r GST Hedrick}
kable(Gst_Hedrick(myData_genind_allMarkersOnly))
#very high differentiation for all loci!

```

```{r locus summary}
kable(locus_table(myData_genind_allMarkersOnly))
```

#Genotypic diversity
Diversity measures incorporate both genotypic richness and abundance. There are three measures of genotypic diversity employed by poppr, the Shannon-Wiener index (H), Stoddart and Taylor’s index (G), and Simpson’s index (lambda). In our example, comparing the diversity of BB to FR shows that H is greater for FR (4.58 vs. 4.4), but G is lower (53.4 vs. 61.7). Thus, our expectation that diversity is lower for FR than BB is rejected in the case of H, which is likely due to the sensitivity of the Shannon-Wiener index to genotypic richness in the uneven sample sizes, and accepted in the case of G. To be fair, the sample size used to calculate these diversity measures is different and is therefore not an appropriate comparison.

For an easier statistic to grasp, we have included the Simpson index, which is simply one minus the sum of squared genotype frequencies. This measure provides an estimation of the probability that two randomly selected genotypes are different and scales from 0 (no genotypes are different) to 1 (all genotypes are different). In the data above, we can see that lambda is just barely higher in BB than FR (0.984 vs. 0.981). Since this might be an artifact of sample size, we can explore a correction of Simpson’s index for sample size by multiplying lambda by N/(N−1). Since R is vectorized, we can do this for all of our populations at once:

```{r diversity}
setPop(myData_genind_allMarkersOnly) <- ~Pop
allMarkersOnly_diversity <- poppr(myData_genind_allMarkersOnly)
N      <- allMarkersOnly_diversity$N      # number of samples
lambda <- allMarkersOnly_diversity$lambda # Simpson's index
lambda_corr <- (N/(N - 1)) * lambda              # Corrected Simpson's index
allMarkersOnly_diversity$lambda_corr <- lambda_corr
allMarkersOnly_diversity
```

```{r diversity auf level cluster}
setPop(myData_genind_withClusters) <- ~cluster
poppr(myData_genind_withClusters)
```

#Genotypic evenness
Evenness is a measure of the distribution of genotype abundances, wherein a population with equally abundant genotypes yields a value equal to 1 and a population dominated by a single genotype is closer to zero --> Value E.5
```{r evenness}
(allMarkersTable <- mlg.table(myData_genind_allMarkersOnly))
```

# Genetic distance
One common way to visualize a genetic distance is with a dendrogram. For this example, we will use the microbov data set (Laloe et al., 2001). This contains information on 704 cattle from both Africa and France over several different breeds. We can create a dendrogram over all 704 samples, but that would be difficult to visualize. For our purposes, let’s take ten random samples and calculate Provesti’s distance, which will return the fraction of the number of differences between samples:

```{r genetic distance}
setPop(myData_genind_allMarkersOnly) <- ~Pop/SamplingYear
data(myData_genind_allMarkersOnly)
set.seed(10)
ten_samples <- sample(nInd(myData_genind_allMarkersOnly), 10)
mic10       <- myData_genind_allMarkersOnly[ten_samples]
(micdist    <- provesti.dist(mic10))
#The above represents the pairwise distances between these 10 samples. We will use this distance matrix to create a neighbor-joining tree.
theTree <- micdist %>%
  nj() %>%    # calculate neighbor-joining tree
  ladderize() # organize branches by clade
plot(theTree)
add.scale.bar(length = 0.05) # add a scale bar showing 5% difference.

set.seed(999)
aboot(mic10, dist = provesti.dist, sample = 200, tree = "nj", cutoff = 50, quiet = TRUE)

#The bootstrap value of 100 on the node separating the French and African samples gives support that the country of origin is a factor in how these breeds are structured. If we wanted to analyze all of the breeds against one another, it would be better to create a bootstrapped dendrogram based on a genetic distance. To do this, we will add 3 stratifications to the microbov data set: Country, Breed, and Species. We will then set the population to Country by Breed, convert the data to a genpop object and then create a tree using aboot with Nei’s genetic distance.

set.seed(999)
myData_genind_allMarkersOnly %>%
  genind2genpop(pop = ~Pop) %>%
  aboot(cutoff = 50, quiet = TRUE, sample = 1000, distance = nei.dist)
#entspricht auch den Ergebnissen von Structure und PCA
```
# K-means hierarchical clustering
A recent study reported that the origin of the potato late blight pathogen Phytophthora infestans lies in Mexico as opposed to South America (Goss et al., 2014). We saw in the previous chapter that South American populations showed signatures of clonal reproduction while Mexican populations showed no evidence rejecting the null hypothesis of random mating. In this section, we will use K-means clustering in combination with bootstrapped dendrograms to see how well this pattern holds up. Clonal populations should have short terminal branch lengths and should cluster according to those branches. Panmictic populations will show no clear pattern. Let’s look at the data:

```{r cluster analysis, eval=FALSE, include=FALSE}
setPop(myData_genind_allMarkersOnly) <- ~Pop
MX <- popsub(myData_genind_allMarkersOnly, "BUR")
MXclust <- find.clusters(MX)
```

```{r MSN}
# KON_msn <- myData_genind_allMarkersOnly %>% setPop(~Pop) %>% popsub("KON")
#args(plot_poppr_msn)
#ls()
# imsn() #accesses html tool to calculate distance networks
#command from the tool:
setPop(myData_genind_allMarkersOnly) <- ~Pop
myData_genind_allMarkersOnly_sub <- popsub(myData_genind_allMarkersOnly, sublist = c("FRE", "FRB", "BUR", "BRU"))
min_span_net <- bruvo.msn(myData_genind_allMarkersOnly_sub, replen = c(1, 2, 3), add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
set.seed(69)
plot_poppr_msn(myData_genind_allMarkersOnly,
               min_span_net,
               inds = "ALL",
               mlg = FALSE,
               gadj = 3,
               nodescale = 10,
               palette = cm.colors,
               cutoff = NULL,
               quantiles = FALSE,
               beforecut = TRUE,
               pop.leg = TRUE,
               size.leg = TRUE,
               scale.leg = TRUE,
               layfun = igraph::layout_nicely)

#with MLGs instead of samples
plot_poppr_msn(myData_genind_allMarkersOnly,
               min_span_net,
               inds = "ALL",
               mlg = TRUE,
               gadj = 3,
               nodescale = 10,
               palette = cm.colors,
               cutoff = NULL,
               quantiles = FALSE,
               beforecut = TRUE,
               pop.leg = TRUE,
               size.leg = TRUE,
               scale.leg = TRUE,
               layfun = igraph::layout_nicely)
```

# AMOVA
In this chapter, we will utilize AMOVA to analyze our populations. AMOVA stands for Analysis of MOlecular VAriance and is a method to detect population differentiation utilizing molecular markers (Excoffier, Smouse & Quattro, 1992). This procedure was initially implemented for DNA haplotypes, but applies to any marker system. The implementation of AMOVA in poppr requires two very basic components: (1) A distance matrix derived from the data and (2) a separate table used to partition the data into different stratifications.
The distance matrix can be calculated using any distance as long as it is euclidean. The distance that is used in the program Arlequin is the opposite of the Kronecker Delta function that counts the number of differences summed over L loci

```{r amova}
#In panmictic populations, we would expect to see most of the variance arise from within samples. If we see that the most of the variance occurs among samples within populations or among populations, then there is evidence that we have some sort of population structure. In the case of clonal organisms, this would help support a hypothesis of clonal reproduction.
allMarkers_amova <- poppr.amova(myData_genind_allMarkersOnly, ~Pop)
allMarkers_amova
allMarkers_amovacc <- poppr.amova(myData_genind_allMarkersOnly, ~Pop, clonecorrect = TRUE)
allMarkers_amovacc
allMarkers_cluster_amova <- poppr.amova(myData_genind_withClusters, ~cluster)
allMarkers_cluster_amova

#Keines davon ergibt das gleiche Resultat wie im Genalex --> Check datasets!

#significance testing
set.seed(1999)
allMarkers_signif   <- randtest(allMarkers_amova, nrepet = 999)
plot(allMarkers_signif)
allMarkers_ccsignif <- randtest(allMarkers_amovacc, nrepet = 999)
plot(allMarkers_ccsignif)
```


# Alleles contributing to differentiation of BUR,FRB and FRE

```{r find private alleles}
#assess if there is alleles in BUR, FRB and FRE that distinguishes them
set.seed(4)
setPop(myData_genind_allMarkersOnly) <- ~Pop
dapc.allMarkers <- dapc(myData_genind_allMarkersOnly, var.contrib = TRUE, scale = FALSE, n.pca = 30, n.da = nPop(myData_genind_allMarkersOnly) - 1)
contrib <- loadingplot(dapc.allMarkers$var.contr, axis = 2, thres = 0.01, lab.jitter = 1, threshold=0.05)

temp    <- seploc(myData_genind_allMarkersOnly)       # seploc {adegenet} creates a list of individual loci.
aest29_1  <- tab(temp[["aest29_1"]])
aest06_1  <- tab(temp[["aest06_1"]]) #allele 217
aest24_1  <- tab(temp[["aest24_1"]]) #allele 295 & 301
aest10_1  <- tab(temp[["aest10_1"]]) #allele 280
aest28_1  <- tab(temp[["aest28_1"]])
aest36_1  <- tab(temp[["aest36_1"]])

# The following two commands find the average allele frequencies per population
setPop(myData_genind_allMarkersOnly) <- ~Pop
(freq29 <- apply(aest29_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))

setPop(myData_genind_allMarkersOnly) <- ~SamplingYear
(freq29 <- apply(aest29_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))

df_29 <- melt(freq29)
names(df_29)[1] <- "SamplingYear"
names(df_29)[2] <- "Allele"
names(df_29)[3] <- "Allele frequencies"

ggplot(df_29,aes(x=SamplingYear,y=`Allele frequencies`)) +
  geom_line(aes(color=Allele)) +
  labs(x="SamplingYear", y="Allele frequency", title="Allele frequencies of aest29 over the years") +
  theme_classic()

setPop(myData_genind_allMarkersOnly) <- ~Pop
(freq36 <- apply(aest36_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))

setPop(myData_genind_allMarkersOnly) <- ~SamplingYear
(freq36 <- apply(aest36_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))

df_36 <- melt(freq36)
names(df_36)[1] <- "SamplingYear"
names(df_36)[2] <- "Allele"
names(df_36)[3] <- "Allele frequencies"

ggplot(df_36,aes(x=SamplingYear,y=`Allele frequencies`)) +
  geom_line(aes(color=Allele)) +
  labs(x="SamplingYear", y="Allele frequency", title="Allele frequencies of aest36 over the years") +
  theme_classic()


setPop(myData_genind_allMarkersOnly) <- ~Pop
(freq06 <- apply(aest06_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))
# The following two commands find the average allele frequencies per population
setPop(myData_genind_allMarkersOnly) <- ~SamplingYear
(freq06 <- apply(aest06_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))
#This allele 217 started appearing in the year 2016, but wasn't found in 2023

df_06 <- melt(freq06)
names(df_06)[1] <- "SamplingYear"
names(df_06)[2] <- "Allele"
names(df_06)[3] <- "Allele frequencies"

ggplot(df_06,aes(x=SamplingYear,y=`Allele frequencies`)) +
  geom_line(aes(color=Allele)) +
  labs(x="SamplingYear", y="Allele frequency", title="Allele frequencies of aest06 over the years") +
  theme_classic()


setPop(myData_genind_allMarkersOnly) <- ~Pop
(freq24 <- apply(aest24_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))
# The following two commands find the average allele frequencies per population
setPop(myData_genind_allMarkersOnly) <- ~SamplingYear
(freq24 <- apply(aest24_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))
# The allele 295 was always present since 2011 but increased during the last years, as can be seen also here graphically:

df_24 <- melt(freq24)
names(df_24)[1] <- "SamplingYear"
names(df_24)[2] <- "Allele"
names(df_24)[3] <- "Allele frequencies"

ggplot(df_24,aes(x=SamplingYear,y=`Allele frequencies`)) +
  geom_line(aes(color=Allele)) +
  labs(x="SamplingYear", y="Allele frequency", title="Allele frequencies of aest24 over the years") +
  theme_classic()

setPop(myData_genind_allMarkersOnly) <- ~Pop
(freq10 <- apply(aest10_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))
# The following two commands find the average allele frequencies per population
setPop(myData_genind_allMarkersOnly) <- ~SamplingYear
(freq10 <- apply(aest10_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))
# the allele 280 was always present, peaking in 2017.

df_10 <- melt(freq10)
names(df_10)[1] <- "SamplingYear"
names(df_10)[2] <- "Allele"
names(df_10)[3] <- "Allele frequencies"

ggplot(df_10,aes(x=SamplingYear,y=`Allele frequencies`)) +
  geom_line(aes(color=Allele)) +
  labs(x="SamplingYear", y="Allele frequency", title="Allele frequencies of aest10 over the years") +
  theme_classic()

setPop(myData_genind_allMarkersOnly) <- ~Pop
(freq28 <- apply(aest28_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))
# The following two commands find the average allele frequencies per population
setPop(myData_genind_allMarkersOnly) <- ~SamplingYear
(freq28 <- apply(aest28_1, 2, function(e) tapply(e, pop(myData_genind_allMarkersOnly), mean, na.rm = TRUE)))
# the allele 280 was always present, peaking in 2017.

df_28 <- melt(freq28)
names(df_28)[1] <- "SamplingYear"
names(df_28)[2] <- "Allele"
names(df_28)[3] <- "Allele frequencies"

ggplot(df_28,aes(x=SamplingYear,y=`Allele frequencies`)) +
  geom_line(aes(color=Allele)) +
  labs(x="SamplingYear", y="Allele frequency", title="Allele frequencies of aest28 over the years") +
  theme_classic()


```

#Isolation by distance (Mantel test)
```{r isolation by distance, eval=FALSE, include=FALSE}
#https://cran.r-project.org/web/packages/graph4lg/vignettes/input_data_processing_1.html
dist_allMarkersOnly <- mat_gen_dist(myData_genind_allMarkersOnly,dist="basic")
coord_dist <- data.frame(coord$Site_1_abrev,coord$LAT,coord$LON)
names(coord_dist)[1] <- "ID"
names(coord_dist)[2] <- "Y" #latitude
names(coord_dist)[3] <- "X" #longitude
mat_geo <- mat_geo_dist(coord_dist, ID="ID",x="X",y="Y", crds_type="polar")

# https://adegenet.r-forge.r-project.org/files/tutorial-basics.pdf

setPop(myData_genind_allMarkersOnly) <- ~Pop
myData_genind_allMarkersOnly@other$xy <- coord_dist[,2:3]
myData_genpop_allMarkersOnly2 <- genind2genpop(myData_genind_allMarkersOnly)
myData_genpop_allMarkersOnly2
Dgen <- dist.genpop(myData_genpop_allMarkersOnly2,method=2)
Dgeo <- dist(myData_genpop_allMarkersOnly2@other$xy)
ibd <- mantel.randtest(Dgen,Dgeo)
ibd
plot(ibd)
plot(Dgeo, Dgen)
#abline(lm(Dgen~log(Dgeo)), col="red",lty=2)

#Hmm... das sieht nicht so befriedigend aus wie im Genalex...
```

```{r isolation by distance V2}
# aus https://bookdown.org/hhwagner1/LandGenCourse_book/WE_5.html

```